MATH/COMP 562 ASSIGNMENT 2

### 2023-06-30 assignment-2/article.tex
Clean up whitespace
First attempt at convex learning

\end{solution}

@@

\end{solution}

@@

\end{solution}

>>>>

\begin{solution}
	Consider three training points in $\mathbb{R}^2$:
	\[
		S = \{((0,0), -1), ((0,1), +1), ((1,0), +1)\}
	\]
	Let's select a vector $w=(0.5,0.5)$. Define the decision boundary as the hyperplane where $w \cdot x = 0.5$. This classifier $w$ has a loss of $L_S(w) = \frac{1}{3}$ because it misclassifies the point ((0,0),-1).

	Now, for a small $\varepsilon > 0$, consider the $\varepsilon$-ball around $w$. For any vector $w'$ within this ball (i.e., $|w - w'| \leq \varepsilon$), we observe that $L_S(w')$ will not be lower than $L_S(w)$. This is because moving the decision boundary within this $\varepsilon$-ball range won't help in correctly classifying the misclassified point. Therefore, the vector $w$ is a local minimum of $L_S$.

	However, consider another vector $w^\prime=(1,0)$ which defines the decision boundary as the hyperplane where $w^\prime \cdot x = 0.5$. This classifier $w^\prime$ correctly classifies all points in the training set S, and hence, its loss $L_S(w^\prime) = 0$. Therefore, we have $L_S(w^\prime) < L_S(w)$, meaning that the vector $w$ is not a global minimum of $L_S$.

	Thus, we have constructed a training sample S for which the 0-1 loss function has a local minimum but not a global minimum. This satisfies all conditions of the problem.
\end{solution}

@@

\begin{solution}
	Given the logistic loss function
	\[\ell(w, (x, y)) = \log(1 + \exp(-y\langle w, x \rangle))\]

	The gradient of the loss function with respect to $w$ is
	\[\nabla \ell(w, (x, y)) = -\frac{y \exp(-y\langle w, x \rangle)}{1+\exp(-y\langle w, x \rangle)}x = -\frac{y}{1+\exp(y\langle w, x \rangle)}x\]

	Its norm can be bounded by $B$ because $\|x\| \leq B$. Thus, the loss function is B-Lipschitz.

	To evaluate smoothness, we look at the second derivative of the loss function with respect to $w$. The Hessian is given by
	\[\nabla^2 \ell(w, (x, y)) = \frac{\exp(y\langle w, x \rangle)}{(1+\exp(y\langle w, x \rangle))^2}xx^T\]

	The spectral norm of the Hessian can be bounded by $B^2$ as $\|x\|^2 \leq B^2$. Thus, the loss function is B²-smooth.

	Therefore, the logistic regression loss function is both B-Lipschitz and B²-smooth.
\end{solution}

@@

	A function $f: \mathbb{R}^d \rightarrow \mathbb{R}$ is Lipschitz if there exists a constant $L > 0$ such that
	\[|f(u) - f(v)| \leq L\|u - v\|\]
	for all $u, v$ in the domain of $f$.

	Here, we are considering the hinge loss function, given by:
	\[\ell(w, (x, y)) = \max\{0, 1 - y\langle w, x \rangle\}\]

	To show that the hinge loss function is Lipschitz continuous, we have to prove that the difference in loss values for two distinct inputs $w$ and $w'$ is at most $R$ times the distance between $w$ and $w'$.

	Suppose $w, w' \in \mathbb{R}^d$ and $(x, y) \in X \times Y$, then:

	\begin{align*}
		| \ell(w, (x, y)) - \ell(w', (x, y)) | & = | \max\{0, 1 - y\langle w, x \rangle\} - \max\{0, 1 - y\langle w', x \rangle\} |                                               \\
		                                       & \leq | y\langle w, x \rangle - y\langle w', x \rangle | \quad \text{(since max function is Lipschitz with Lipschitz constant 1)} \\
		                                       & = | y\langle w - w', x \rangle |                                                                                                 \\
		                                       & \leq \| y(x) \|_2 \| w - w' \|_2 \quad \text{(by Cauchy-Schwarz inequality)}                                                     \\
		                                       & \leq R \| w - w' \|_2 \quad \text{(since } \| x \|_2 \leq R \text{)}
	\end{align*}

	Therefore, the hinge loss function $\ell$ is $R$-Lipschitz.
\end{solution}
